---
output: pdf_document
---
```{r Libraries, echo=FALSE}
library(ggplot2)

```
```{r Function plot with fill, echo=FALSE}
#' Draw Normal Distribution Density with an area shaded in.
#'
#' @param lb Lower bound of the shaded area. Use \code{-Inf} for a left tail.
#' @param ub Upper bound of the shaded area. Use \code{Inf} for a right tail.
#' @param mean Mean of the normal distribution
#' @param sd Standard deviation of the normal distribution
#' @param limits Lower and upper bounds on the x-axis of the area displayed.
#' @return ggplot object.
#' @examples
#' # Standard normal with upper 2.5% tail shaded
#' normal_prob_area_plot(2, Inf)
#' # Standard normal with lower 2.5% tail shaded
#' normal_prob_area_plot(-Inf, 2)
#' # standard normal with middle 68% shaded.
#' normal_prob_area_plot(-1, 1)
plot_norm <- function(lb, 
                      ub, 
                      lb1 = "a",
                      ub1 = NA,
                      lb2 = "a",
                      ub2 = NA,
                      lb3 = "a",
                      ub3 = NA,
                      mean = 0, 
                      sd = 1, 
                      limits = c(mean - 3 * sd, mean + 3 * sd), 
                      col = "gray78",
                      col1 = "gray68",
                      col2 = "gray58",
                      col3 = "gray48",
                      alpha = .4) {
    x <- seq(limits[1], limits[2], length.out = 100)
    xmin <- max(lb, limits[1])
    xmax <- min(ub, limits[2])
    areax <- seq(xmin, xmax, length.out = 100)
    area <- data.frame(x = areax, ymin = 0, ymax = dnorm(areax, mean = mean, sd = sd))
    g <- ggplot()
    g <- g + geom_line(data.frame(x = x, 
                                  y = dnorm(x, mean = mean, sd = sd)),
                       mapping = aes(x = x, y = y))
    g <- g + geom_ribbon(data = area, 
                         mapping = aes(x = x, ymin = ymin, ymax = ymax), 
                         fill = col, 
                         alpha = alpha)
    for(i in 1:3) {
        lb_i <- get(paste("lb", i, sep = ""))
        ub_i <- get(paste("ub", i, sep = ""))
        col_i <- get(paste("col", i, sep = ""))
        if(lb_i != "a") {
            xmin <- max(lb_i, limits[1])
            xmax <- min(ub_i, limits[2])
            g <- g + geom_ribbon(data = area, 
                                 mapping = aes(x = x, ymin = ymin, ymax = ymax), 
                                 fill = col_i, 
                                 alpha = alpha)
            }
           }
    
    g <- g + scale_x_continuous(limits = limits)
    g
    }
```
## DEFINITIONS ##
**POPULATION** : Entire aggregate of individuals or items from which SAMPLES are drawn.   

**SAMPLE** : A set of individuals or items selected from a PARENT POPULATION so that properties or parametres of the population may be estimated.   
**Sampling distribution of the means** : distribution of the mean of samples of particular size.
**Sampling distribution of the variance** : distribution of the variance of samples.

**STATISTIC** : any function of sample data, containing no unknown parametres, such as mean, median, variance or standard deviation.    
**Population's statistics** are generally indicated with Greek letters ($\mu$, $\sigma$, etc.).  
**Sample's statistics** are generally indicated with Latin letters (m, s, etc.).  

**RANDOM VARIABLE** : a variable which takes values in certain range with probabilities specified by a PROBABILITY DENSITY FUNCTION or PROBABILITY MASS FUNCTION (ex. if we express head or tail as 0 or 1, the toss of a coin is a random variable).  

**EXPECTED VALUE** : or **MEAN** of a random variable or a function of a variable  E[X] or $\bar{X}$ is

1. $\sum_{i=1}^{n}x_ip(x_i)$ for a DISCRETE VARIABLE;  
2. $\int x f(x)dx$ for a CONTINUOUS VARIABLE.  

It has the following properties:  
E[$\lambda$ X] = $\lambda$ E[X]  
E[X $\pm$ Y] = E[X] $\pm$ E[Y]  
E[$X^2$] = $E[X]^2$ + Var(X)  
E[XY] = E[X]E[Y] + cov(X,Y)  

**VARIANCE** : a measure of dispersion  
   
* *Theoretical variance* $$Var(X) = \sum_{i=1}^{n} \frac{(x_i - \bar{x})^2}{n} = E[X^2] - E[X]^2$$  
   
* *Sample variance* $$s^2 = \frac{\sigma^2}{n}$$   

**STANDARD DEVIATION** is the square root of the variation and has the same unit of the random variable 

* *Theoretical standard deviation* $$\sigma = \sqrt{Var(X)}$$ 

* *Sample standard deviation* = *Standard error* $$s = \frac{\sigma}{\sqrt{n}}$$

$\pagebreak$

**PROBABILITY MASS FUNCTION (PMF)** : the distribution of the probability of the different values of a discrete random variable X.
If the variable X takes values $x_1, ..., x_n$ with probabilities $p(x_1), ..., p(x_n)$, than:  
1. $\sum_{i} p(x_i) = 1$  
2. $p(x_i)\leq 0$ $\forall i$

**PROBABILITY DENSITY FUNCTION (PDF)** : the function f(x) of a continuous variable X such that:  
1. the probability that X lies between $x_0$ and $x_1$ is $\int_\mathrm{x_0}^\mathrm{x_1} \mathrm{f(x)} \mathrm{d}x$   
2. the cumulative probability of the whole range of x is equal to 1 : $\int_\mathrm{x} \mathrm{f(x)} \mathrm{d}x = 1$  
3. the probability is always greater than 0 : $f(x) \leq 0 \forall x$ 

**PROBABILITY** : a measure of the relative frequency or likelyhood of occurrence of an event. Values are deived from a **theoretical distribution** or from **observations**.  

- **P(x)** is the probability of the event x.     
- $0 \leq P(x) \leq 1$  
- **Discrete variables** : $\frac {number of required outcomes}{total number of possible outcomes}$   
- **Continuous variables** : The relevant area under the graph of its probability density function f(x) $$ P(x) = \int_\mathrm{x_0}^\mathrm{x_1} \mathrm{f(x)} \mathrm{d}x  $$

```{r, echo=FALSE}
g <- plot_norm(lb = 0, ub = 0.8, col = "blue")
g <- g + annotate("text", x = -1.3, y = 0.3, parse = TRUE, label =" frac( 1, sqrt( 2 * pi)) * e ^ {-x ^ 2 / 2}", size = 4)
g <- g + annotate("text", x = -1.9, y = 0.3, parse = FALSE, label ="f(x) = ", size = 4)
g <- g + annotate("text", x = -0.2, y = 0, label="x[0]", parse=TRUE)
g <- g + annotate("text", x = 1, y = 0, label="x[1]", parse=TRUE)
g <- g + annotate("text", x = 0.3, y = 0.17, label = "P(x)")
g
```     

$\pagebreak$

**QUANTILE** : general name for the values of a variable which divides its distribution into equal groups
```{r}
qnorm(.95)
```   

```{r, echo=FALSE,fig.width=5, fig.height=3.5}
g <- plot_norm(lb = -Inf, ub = qnorm(.95), col = "blue")
g <- g + annotate("text", x = round(qnorm(.95), 2), y = -0.015, label= paste("x = ",round(qnorm(.95), 2)))
g <- g + annotate("text", x = 0, y = 0.04, label = "95%")
g <- g + annotate("text", x = 2, y = 0.03, label = "5%")
g
```

**CUMULATIVE DISTRIBUTION FUNCTION (CDF)** : the function F(x) which gives the cumulative frequency

$\pagebreak$

## DISTRIBUTIONS ##

**BERNOULLI DISTRIBUTION** : a type of binomial distribution when the random variables take only values 0 or 1 with probability p and 1-p, respectively  

* *Probability mass function* : $P(X = x_1) = p^{x_1}(1-p)^{1-x_1}$ with X = [0,1]  
     
* *Expected value* : $\mu = p$  

* *Variance* : $Var(X) = p(1-p)$   

**BINOMIAL** : is obtained as the sum of a bunch of iid bernoulli random variables (ex. number of heads on a biased coin). Let $x_1, ..., X_n$ be an iid Bernoulli with probability p, then $$X = \sum{i=1}^{n} x_i$$ is a *binomial random variable* with mass function $$P(X = x_i) = \binom{n}{k}p^{x_1}(1-p)^{1-x_1}$$  

**NORMAL DISTRIBUTION** Gaussian distribution with mean $\mu$ and variance $\sigma^2$  
$$X \sim N(\mu,\sigma^2)$$  

* *Probability Mass Function* : $$f(x) = (2\pi\sigma^2)^{-1/2}e^{-\frac{(x-\mu)^2}{2\sigma^2}}$$

**STANDARD NORMAL DISTRIBUTION** : is the specific case of a normal distribution with $\mu = 0$ and $\sigma = 1$  
$$X \sim N(0,1)$$ 
with probability mass function $$f(x) = \frac{e^{-1/2x^2}}{\sqrt{2\pi}}$$


